Thermodynamic fragments in magic state theories

### 2021-01-18 magic_paper.tex
Outline for thermal n-copy lorenz curve generating functions, need to understand maximum

Consider a system with Hamiltonian $H$ and in contact with a thermal bath of temperature $\beta^{-1}$.
We denote the thermal state at that temperature by $\gamma \equiv \gamma_\beta = \frac{e^{-\beta H}}{\Z_\beta}$, where $\Z \equiv \Z_\beta = \tr[e^{-\beta H}]$.
As long as the thermal state is free, it is associated with a set of free, thermal operations and the dynamics of the system can be described within the fragment $\O_{\gamma}$.

We want to focus on qutrit interconversions and extract a temperature-dependent distillation bound.
To this end, consider the qutrit Hamiltonian
\begin{equation}
	H = E_0 \ketbra{0} + E_1 \ketbra{1} + E_2 \ketbra{2},
\end{equation}
where the eigenbasis coincides with the computational ($Z$) basis.
The thermal state associated with this Hamiltonian is
\begin{align}
	\gamma &= \frac{e^{-\beta H}}{\Z} = \frac{e^{-\beta \sum\limits_{k=0}^2 E_k \ketbra{k}}}{\Z} \\
	&= \frac{e^{-\beta E_0}}{\Z}\ketbra{0} + \frac{e^{-\beta E_1}}{\Z}\ketbra{1} + \frac{e^{-\beta E_2}}{\Z}\ketbra{2},
\end{align}
where $\Z = e^{-\beta E_0} + e^{-\beta E_1} + e^{-\beta E_2}$.

The Wigner distribution of state $\gamma$ can be seen as the ensemble average of the distributions of the computational basis states,
\begin{align}
	\W[\bmx]{\gamma} &= \sum\limits_{k=0}^2 \frac{e^{-\beta E_k}}{\Z}\W[\bmx]{\ketbra{k}} \\
	&= \sum\limits_{k=0}^2 \frac{e^{-\beta E_k}}{\Z} \delta_{x_0, k} = \frac{e^{-\beta E_{x_0}}}{d\Z},
\end{align}
for all $\bmx \in \pd$. 
All Wigner components are strictly positive, therefore the pre-order $\prec_{\gamma_\beta}$ is well defined.

Consider the simple energy spectrum 
\begin{equation}
	(E_0, E_1, E_2) = (1,0,0).
\end{equation}
Then, $Z = 2 + e^{-\beta}$.
Furthermore, for clarity, let $\chi \coloneqq \frac{1}{6}-\frac{1}{18}\epsilon$ be the positive and $-\psi \coloneqq -\frac{1}{3} + \frac{4}{9}\epsilon$ the negative Wigner components of $\noisys(\epsilon)$.
The 1-copy Wigner distributions in this scenario are depicted in~\cref{tab:thermal_w}.
\begin{table}[h]
  \def\arraystretch{2}
  \centering
  \begin{tabular}{c|c|c|c}
    $\bmx = (x_0, x_1)$ & $W_{\noisys}(\bmx)$ & $W_{\gamma}(\bmx)$ & $\tilde{\bmw}$ \\
    \hline
    \makecell{$(0,0)$ \\ \vdots \\ $(0,2)$} & $\displaystyle \chi$ & $\displaystyle \frac{e^{-\beta}}{3Z}$ & $\displaystyle 3Z \chi e^{\beta}$ \\
    \hline
    \makecell{$(1,0)$ \\ $\vdots$ \\ $(2,1)$} & $\displaystyle \chi$ & $\displaystyle \frac{1}{3Z}$ & $\displaystyle 3Z \chi$ \\
    \hline
    $(2,2)$ & $\displaystyle -\psi$ & $\displaystyle \frac{1}{3Z}$ & $\displaystyle 3Z (-\psi)$ \\
    \hline
  \end{tabular}
  \caption{Wigner components $W_{\gamma}(\bmx)$, $W_{\noisys}(\bmx)$ and the associated slope $\tilde{\bmw}$ listed in order of decreasing slope (holds for all temperatures $\beta^{-1}$).}
  \label{tab:thermal_w}
\end{table}

@@

Notice that the slope of the $n$-copy Lorenz curve can be expressed as a self-convolution of the 1-copy Lorenz curve slope $n$ times,
\begin{equation}
	\tilde{\bmw\too{n}} = \frac{W_{\noisys\too{n}}(\bmx)}{W_{\gamma\too{n}}(\bmx)} = \left( \frac{W_{\noisys}(\bmx)}{W_{\gamma}(\bmx)} \right)\too{n} = \tilde{\bmw}\too{n}.
\end{equation}

According to~\cref{def:lc}, to construct the $n$-copy Lorenz curve, we need to sort $\tilde{\bmw\too{n}}$ in decreasing order and take the cumulative sum of the corresponding components $W_{\noisys\too{n}}(\bmx)$ and $W_{\gamma\too{n}}(\bmx)$.

@@

For a qutrit, let the energy spectrum be $E_0 \geq E_1 \geq E_2$. 
Then, we can explicitly write the thermal Wigner distribution as
\begin{equation}
	\W{\gamma_\beta} = \frac{1}{3\Z_\beta} \big( \overbrace{e^{-\beta E_0}}^3, \overbrace{e^{-\beta E_1}}^3, \overbrace{e^{-\beta E_2}}^3 \big),
\end{equation}
which is sorted in decreasing order.
We choose the Strange state where the negativity is in the first component of its Wigner distribution
\begin{equation}\label{eq:wsu}
	\W{\noisys(\epsilon)} = \Bigg( \overbrace{-\frac{1}{3} + \frac{4}{9}\epsilon}^1, \overbrace{\frac{1}{6} - \frac{1}{18}\epsilon}^8\Bigg),
\end{equation}

However, higher copies are not sorted, for example
\begin{equation}
	\begin{split}
	\W{\gamma_\beta\too{2}} = \frac{1}{9\Z_\beta^2} \Bigg( &\overbrace{\overbrace{e^{-\beta 2E_0}}^3, \overbrace{e^{-\beta (E_0 + E_1)}}^3, \overbrace{e^{-\beta (E_0 + E_2)}}^3}^3, \\
	 &\overbrace{\overbrace{e^{-\beta (E_1 + E_0)}}^3, \overbrace{e^{-\beta 2E_1}}^3, \overbrace{e^{-\beta (E_1 + E_2)}}^3}^3, \\
	 &\overbrace{\overbrace{e^{-\beta (E_2 + E_0)}}^3, \overbrace{e^{-\beta (E_2 + E_1)}}^3, \overbrace{e^{-\beta 2E_2}}^3}^3 \Bigg).
	\end{split}
\end{equation}
\ddd{[Definitely impressed by this Latexing...but perhaps just give the multiplicities and spectrum values. I.e. ``(blah) with multiplicity (blah) and ...'']} \nick{So the problem is that the exact position of each component matters, e.g. thermal state component $e^{-\beta 2E_0}$ corresponds to a positive Strange state component in one position and a negative Strange state component in another.} 
For the Strange state we have
\begin{equation}
	\begin{split}
	\W{\noisys(\epsilon)\too{2}} = \Bigg( &\overbrace{\left( -\frac{1}{3} + \frac{4}{9}\epsilon \right)^2}^1, \overbrace{ \left( -\frac{1}{3} + \frac{4}{9}\epsilon \right) \left( \frac{1}{6} - \frac{1}{18}\epsilon \right)}^{8}, \\
	&\overbrace{\overbrace{ \left( -\frac{1}{3} + \frac{4}{9}\epsilon \right) \left( \frac{1}{6} - \frac{1}{18}\epsilon \right)}^1, \overbrace{\left( \frac{1}{6} - \frac{1}{18}\epsilon \right)^2}^8}^8,  \Bigg).
	\end{split}
\end{equation}
\nick{We need to sort the distributions $\W{\gamma_\beta\too{2}}, \W{\noisys(\epsilon)\too{2}}$, so that the vector
\begin{equation}
	\tilde{\bm{w}} \coloneqq \left(\frac{(\W{\noisys(\epsilon)\too{2}})_1}{(\W{\sigma})_1}, \dots, \frac{(\W{\rho})_{81}}{(\W{\sigma})_{81}} \right)^T,
\end{equation}
is in decreasing order, i.e. $\tilde{\bm{w}} = \tilde{\bm{w}}^\downarrow$.
Then, we need to generalise for any number of copies $n$}

>>>>

Consider a system with Hamiltonian $H$ and in contact with a thermal bath of temperature $\beta^{-1}$.
We denote the thermal state at that temperature by $\gamma \equiv \gamma_\beta = \frac{e^{-\beta H}}{E_\beta}$, where $E \equiv E_\beta = \tr[e^{-\beta H}]$.
As long as the thermal state is free, it is associated with a set of free, thermal operations and the dynamics of the system can be described within the fragment $\O_{\gamma}$.

We want to focus on qutrit interconversions and extract a temperature-dependent distillation bound.
To this end, consider the qutrit Hamiltonian
\begin{equation}
	H = E_0 \ketbra{0} + E_1 \ketbra{1} + E_2 \ketbra{2},
\end{equation}
where the eigenbasis coincides with the computational basis.
The thermal state associated with this Hamiltonian is
\begin{align}
	\gamma &= \frac{e^{-\beta H}}{E} = \frac{e^{-\beta \sum\limits_{k=0}^2 E_k \ketbra{k}}}{E} \\
	&= \frac{e^{-\beta E_0}}{E}\ketbra{0} + \frac{e^{-\beta E_1}}{E}\ketbra{1} + \frac{e^{-\beta E_2}}{E}\ketbra{2},
\end{align}
where $E = e^{-\beta E_0} + e^{-\beta E_1} + e^{-\beta E_2}$.

The Wigner distribution of state $\gamma$ can be seen as the ensemble average of the distributions of the computational basis states,
\begin{align}
	\W[\bmx]{\gamma} &= \sum\limits_{k=0}^2 \frac{e^{-\beta E_k}}{E}\W[\bmx]{\ketbra{k}} \\
	&= \sum\limits_{k=0}^2 \frac{e^{-\beta E_k}}{E} \delta_{x_0, k} = \frac{e^{-\beta E_{x_0}}}{dE},
\end{align}
for all $\bmx \in \pd$. 
All Wigner components are strictly positive, therefore the pre-order $\prec_{\gamma_\beta}$ is well defined.

Consider the simple energy spectrum 
\begin{equation}
	(E_0, E_1, E_2) = (1,0,0).
\end{equation}
Then, $Z = 2 + e^{-\beta}$.
Furthermore, for clarity, let $\chi \coloneqq \frac{1}{6}-\frac{1}{18}\epsilon$ be the positive and $-\nu \coloneqq -\frac{1}{3} + \frac{4}{9}\epsilon$ the negative Wigner components of $\noisys(\epsilon)$.
The 1-copy Wigner distributions in this scenario are depicted in~\cref{tab:thermal_w}.
\begin{table}[h]
  \def\arraystretch{2}
  \centering
  \begin{tabular}{c|c|c|c}
    $\bmx = (x_0, x_1)$ & $W_{\noisys}(\bmx)$ & $W_{\gamma}(\bmx)$ & $\tilde{\bmw}$ \\
    \hline
    \makecell{$(0,0)$ \\ \vdots \\ $(0,2)$} & $\displaystyle \chi$ & $\displaystyle \frac{e^{-\beta}}{3Z}$ & $\displaystyle 3Z \chi e^{\beta}$ \\
    \hline
    \makecell{$(1,0)$ \\ $\vdots$ \\ $(2,1)$} & $\displaystyle \chi$ & $\displaystyle \frac{1}{3Z}$ & $\displaystyle 3Z \chi$ \\
    \hline
    $(2,2)$ & $\displaystyle -\nu$ & $\displaystyle \frac{1}{3Z}$ & $\displaystyle 3Z (-\nu)$ \\
    \hline
  \end{tabular}
  \caption{Wigner components $W_{\gamma}(\bmx)$, $W_{\noisys}(\bmx)$ and the associated slope $\tilde{\bmw}$ listed in order of decreasing slope (holds for all temperatures $\beta^{-1}$).}
  \label{tab:thermal_w}
\end{table}

@@

Notice that the slope of the $n$-copy Lorenz curve can be expressed as a self-convolution of the 1-copy Lorenz curve slope $n$ times,
\begin{equation}
	\widetilde{\bmw\too{n}} = \frac{W_{\noisys\too{n}}(\bmx)}{W_{\gamma\too{n}}(\bmx)} = \left( \frac{W_{\noisys}(\bmx)}{W_{\gamma}(\bmx)} \right)\too{n} = \widetilde{\bmw}\too{n}.
\end{equation}

According to~\cref{def:lc}, to construct the $n$-copy Lorenz curve, we need to sort $\widetilde{\bmw\too{n}}$ in decreasing order and take the cumulative sum of the corresponding components $W_{\noisys\too{n}}(\bmx)$ and $W_{\gamma\too{n}}(\bmx)$.

We can therefore construst a trinomial generating function that returns the values and the multiplicities of the distinct Wigner components for a general $n$-copy $\epsilon$-noisy Strange state in a $\gamma_\beta$-fragment.
\begin{equation}
	\tilde{w}(X,Y) \coloneqq (3Z)^n (-\nu + 5\chi X + 3\chi e^\beta Y)^n
\end{equation}
There are $\frac{(n+1)(n+2)}{2}$ distinct $X^i Y^j$ terms ($i,j = 0,\dots,n$ with $i+j \leq n$) with the coefficients of these terms representing the multiplicities and values of the different ratios.

We can inductively show that the ``start of the Lorenz curve'' corresponds to $\nu^n$ for even $n$ and $\nu^(n-1) \chi e^\beta$ for 
\begin{equation}
	\beta \leq \ln{\left(8-\frac{18}{3-\epsilon}\right)}
\end{equation}
\nick{Not good enough to restrict to the start of the curve (show why)}

\nick{coordinate generating functions}

@@

For a qutrit, let the energy spectrum be $E_0 \geq E_1 \geq E_2$. 
Then, we can explicitly write the thermal Wigner distribution as
\begin{equation}
	\W{\gamma_\beta} = \frac{1}{3E_\beta} \big( \overbrace{e^{-\beta E_0}}^3, \overbrace{e^{-\beta E_1}}^3, \overbrace{e^{-\beta E_2}}^3 \big),
\end{equation}
which is sorted in decreasing order.
We choose the Strange state where the negativity is in the first component of its Wigner distribution
\begin{equation}\label{eq:wsu}
	\W{\noisys(\epsilon)} = \Bigg( \overbrace{-\frac{1}{3} + \frac{4}{9}\epsilon}^1, \overbrace{\frac{1}{6} - \frac{1}{18}\epsilon}^8\Bigg),
\end{equation}

However, higher copies are not sorted, for example
\begin{equation}
	\begin{split}
	\W{\gamma_\beta\too{2}} = \frac{1}{9E_\beta^2} \Bigg( &\overbrace{\overbrace{e^{-\beta 2E_0}}^3, \overbrace{e^{-\beta (E_0 + E_1)}}^3, \overbrace{e^{-\beta (E_0 + E_2)}}^3}^3, \\
	 &\overbrace{\overbrace{e^{-\beta (E_1 + E_0)}}^3, \overbrace{e^{-\beta 2E_1}}^3, \overbrace{e^{-\beta (E_1 + E_2)}}^3}^3, \\
	 &\overbrace{\overbrace{e^{-\beta (E_2 + E_0)}}^3, \overbrace{e^{-\beta (E_2 + E_1)}}^3, \overbrace{e^{-\beta 2E_2}}^3}^3 \Bigg).
	\end{split}
\end{equation}
\ddd{[Definitely impressed by this Latexing...but perhaps just give the multiplicities and spectrum values. I.e. ``(blah) with multiplicity (blah) and ...'']} \nick{So the problem is that the exact position of each component matters, e.g. thermal state component $e^{-\beta 2E_0}$ corresponds to a positive Strange state component in one position and a negative Strange state component in another.} 
For the Strange state we have
\begin{equation}
	\begin{split}
	\W{\noisys(\epsilon)\too{2}} = \Bigg( &\overbrace{\left( -\frac{1}{3} + \frac{4}{9}\epsilon \right)^2}^1, \overbrace{ \left( -\frac{1}{3} + \frac{4}{9}\epsilon \right) \left( \frac{1}{6} - \frac{1}{18}\epsilon \right)}^{8}, \\
	&\overbrace{\overbrace{ \left( -\frac{1}{3} + \frac{4}{9}\epsilon \right) \left( \frac{1}{6} - \frac{1}{18}\epsilon \right)}^1, \overbrace{\left( \frac{1}{6} - \frac{1}{18}\epsilon \right)^2}^8}^8,  \Bigg).
	\end{split}
\end{equation}
\nick{We need to sort the distributions $\W{\gamma_\beta\too{2}}, \W{\noisys(\epsilon)\too{2}}$, so that the vector
\begin{equation}
	\tilde{\bm{w}} \coloneqq \left(\frac{(\W{\noisys(\epsilon)\too{2}})_1}{(\W{\sigma})_1}, \dots, \frac{(\W{\rho})_{81}}{(\W{\sigma})_{81}} \right)^T,
\end{equation}
is in decreasing order, i.e. $\tilde{\bm{w}} = \tilde{\bm{w}}^\downarrow$.
Then, we need to generalise for any number of copies $n$}
